Representation theory of Finite Groups

### 2016-11-17 Lecture note - Group Theory in Physics/serge lang/s-lang.tex
Read a bit about characters in S. Lang's language. I should learn more about modules to understand the idea of simple module and simple ring.

Suppose now that the modules $E$,$F$ are free and finite dimensional
over $R$ (recall that a module is free when it has a basis). Let
$\rho$ be a representation of $G$ on $E$, $M$ be the matrix of
$\rho(\sigma)$ with respect to a basis, $M^\vee$ be the matrix of
$\rho^\vee (\sigma)$ with respect to the dual basis. Then
\begin{align*}
    [\sigma] f: x &\mapsto f(\sigma^{-1} x) \\
    x^i &\mapsto \sum_{ij} f_i(M^{-1})^i_j x^j \\
    \Longrightarrow [\sigma]: f &\mapsto f\sigma^{-1} \\
    f_j &\mapsto \sum_i f_i(M^{-1})^i_j \\
    \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) &\mapsto (M^{-1})^T  \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) 
\end{align*}
Hence
\begin{equation}
    M^\vee = (M^{-1})^T
\end{equation}

@@

@@

\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

Suppose now that the modules $E$,$F$ are free and finite dimensional
over $R$ (recall that a module is free when it has a basis). Let
$\rho$ be a representation of $G$ on $E$, $M$ be the matrix of
$\rho(\sigma)$ with respect to a basis, $M^\vee$ be the matrix of
$\rho^\vee (\sigma)$ with respect to the dual basis. Then
\begin{align*}
    [\sigma] f: x &\mapsto f(\sigma^{-1} x) \\
    x^i &\mapsto \sum_{ij} f_i(M^{-1})^i_j x^j \\
    \Longrightarrow [\sigma]: f &\mapsto f\sigma^{-1} \\
    f_j &\mapsto \sum_i f_i(M^{-1})^i_j \\
    \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) &\mapsto (M^{-1})^T  \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) 
\end{align*}
Hence
\begin{equation}
    \label{eq:M-dual-M-1T}
    M^\vee = (M^{-1})^T
\end{equation}

@@

\section{Characters}
\label{sec:Characters}
Assuming in this section that, $G$: a finite group. $k$: a field of
characteristic not dividing $|G|$. $E,F$: finite dimensional $k$-spaces.
$n=|G|$.

\begin{defi}[Character $\chi$]
\nomenclature{Character $\chi$}{\nomrefpage.}
    The character $\chi$ of a representation $\rho:k[G]\to \End_k(E)$
    is defined as
    \begin{align}
        \chi: k[G] &\to k \\
        \alpha &\mapsto \Tr(\rho(\alpha))
    \end{align}
    We sometimes write it as $\chi_\rho$ or $\chi_E$.
\end{defi}
The \nomen{trivial character} is sometimes denoted as \nomen{$\chi_0$}
or \nomen{$1_G$}.

\begin{remark}
By the $k$-linearity of trace function, we observe that characters are
determined by its values on $G$.
\end{remark}

\begin{defi}[Isomorphic Representations]
\nomenclature{Isomorphic Representations}{\nomrefpage.}
    Two representations $\rho,\phi$ of $G$ on spaces $E,F$ are
    isomorphic if and only if  there is a $G$-isomorphism between $E$
    and $F$.
\end{defi}
\begin{equation} \begin{tikzcd}[column sep=small]
    & G\ar[dl,"\rho" left=3]\ar[dr,"\phi"] &  \, \\
        E \ar[rr,"\rho" below=3] &
            \ar[u,phantom, "\#"] 
                & F
\end{tikzcd} \end{equation}
\begin{align}
    f(\rho(\sigma)) &= \phi(\sigma) f, \, \forall \sigma\in G
\end{align}
In matrix notation, 
\begin{equation}
    F \rho_\sigma = \phi_\sigma F,\, \forall\sigma\in G
\end{equation}
\textbf{Note}: In everything that follows, we are interested only in
isomorphism classes of representations.

If $E,F$ are $G$-spaces, then their direct sum $E\oplus F$ is also a
$G$-space, the operation of $G$ being componentwise. If $x\oplus y\in
E\oplus F$ with $x\in E,\,y\in F$, then $\sigma(x\oplus y)=\sigma
x\oplus \sigma y$.

Similarly, the tensor product $E\otimes_k F=E\otimes F$ is a
$G$-space, the operation $G$ being given by $\sigma(x\otimes y)=\sigma
x\otimes \sigma y$.

\begin{prop}
    \label{prop:character-ring}
    If $e,F$ are $G$-spaces, then
    \begin{align}
        \chi_E +\chi_F &= \chi_{E\oplus F} \\
        \chi_E\chi_F   &= \chi_{E\otimes F}
    \end{align}
    If $\chi^\vee$ denotes the character of the dual representation on
    $E^\vee$, then
    \begin{align}
        \chi^\vee(\sigma) = \chi(\sigma^{-1})
    \end{align}
    In particular, when $k=\mathbb{C}$, then
    \begin{equation}
        \chi^\vee(\sigma) = \overline{\chi(\sigma)}
    \end{equation}
\end{prop}
\begin{proof}
    The first relation is obvious.

    The second relation holds from the following observation,
    \begin{align*}
        \sigma(v_i \otimes w_j) = \sigma v_i \otimes \sigma w_j =
        A_{i\nu} v_v \otimes B_{j\mu} w_\mu
    \end{align*}
    So its character is
    \begin{equation*}
        \chi_{E\otimes F}(\sigma) = A_{ii} B_{jj} =
        \chi_E(\sigma)\chi_F(\sigma)
    \end{equation*}
    The third relation follows from that the matrix of dual
    representation is the matrix $(M^{-1})^T$, see equation
    \ref{eq:M-dual-M-1T}.

    The last relation will be proved in corollary % TODO
\end{proof}

We now extend the definition of \nomen{character} to a function of
$G$, which can be written in a linear combination of characters of
representations with arbitrary integer coefficients. Then, the
previous characters are very special. Such characters associated with
representations will be called \nomen{effective characters}. Note
that everything we have defined of course depends on the field $k$,
and we shall add \textbf{over} $\mathbf{k}$ to ouirexpressions if we
need to specify the field $k$.

\begin{remark}
    By proposition \ref{prop:character-ring}, all the characters form
    a ring structure.
\end{remark}

By a \nomen{simple} or \nomen{irreducible character} of $G$ one means
the character of a simple representation, i.e., the character
associated with a simple $k[G]$-module.

@@

\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.

\printnomenclature
\end{document}
